---
title: "R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---

loading required library and data 
```{r}
library(tibble)
library(readr)
library(keras)
jena_climate_2009_2016 <- read_csv("jena_climate_2009_2016.csv")

# as data is large we willuse only small portion of data
data <- jena_climate_2009_2016[1:20000,]
data <- data.matrix(data[,-1])


# creating training data for 15000 observation

train_data <- data[1:15000,]

# scaling data on mean and std. dev.
mean <- apply(train_data, 2, mean)
std <- apply(train_data, 2, sd)

data <- scale(data, center = mean, scale = std)

```

 Generator yielding timeseries samples and their targets
```{r}
generator <- function(data, lookback, delay, min_index, max_index,
                      shuffle = FALSE, batch_size = 128, step = 6) {
  if (is.null(max_index))
    max_index <- nrow(data) - delay - 1
  i <- min_index + lookback
  function() {
    if (shuffle) {
      rows <- sample(c((min_index+lookback):max_index), size = batch_size)
    } else {
      if (i + batch_size >= max_index)
        i <<- min_index + lookback
      rows <- c(i:min(i+batch_size, max_index))
      i <<- i + length(rows)
    }
    samples <- array(0, dim = c(length(rows),
                                lookback / step,
                                dim(data)[[-1]]))
    targets <- array(0, dim = c(length(rows)))
    for (j in 1:length(rows)) {
      indices <- seq(rows[[j]] - lookback, rows[[j]],
                     length.out = dim(samples)[[2]])
      samples[j,,] <- data[indices,]
      targets[[j]] <- data[rows[[j]] + delay,2]
    }
    list(samples, targets)
  }
}

```
 
 Preparing the training, validation, and test generators
```{r}
lookback <- 1440
step <- 6
delay <- 144
batch_size <- 128
train_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 1,
  max_index = 15000,
  shuffle = TRUE,
  step = step,
  batch_size = batch_size
)
val_gen = generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 15001,
  max_index = 17000,
  step = step,
  batch_size = batch_size
)
test_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 17001,
  max_index = 19000,
  step = step,
  batch_size = batch_size
)

# to check how many steps to draw from val_gen in order to see the entire validation set

val_steps <- (17000 - 15001 - lookback) / batch_size

# to check howmany steps to draw from test_gen in order to see the entire test set
test_steps <- (19000 - 17001 - lookback) / batch_size

```
 
 Computing the common-sense baseline MAE
```{r}
evaluate_naive_method <- function() {
  batch_maes <- c()
  for (step in 1:val_steps) {
    c(samples, targets) %<-% val_gen()
    preds <- samples[,dim(samples)[[2]],2]
    mae <- mean(abs(preds - targets))
    batch_maes <- c(batch_maes, mae)
  }
  print(mean(batch_maes))
}
(evaluate_naive_method())


```
 
 
 Training and evaluating a densely connected model
```{r}
model <- keras_model_sequential() %>%
  layer_flatten(input_shape = c(lookback / step, dim(data)[-1])) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)

plot(history)
```
 
 
 
  Training and evaluating a model with layer_gru by chnaging numer of units 
```{r echo=TRUE}
model <- keras_model_sequential() %>% 
  layer_gru(units = 64, 
            dropout = 0.1, 
            recurrent_dropout = 0.5,
            return_sequences = TRUE,
            input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_gru(units = 128, activation = "relu",
            dropout = 0.1,
            recurrent_dropout = 0.5) %>% 
  layer_dense(units = 1)
model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)
history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 50,
  epochs = 5,
  validation_data = val_gen,
  validation_steps = val_steps
)

plot(history)


```
  
```{r echo=TRUE}
test <- model %>% fit_generator(
  test_gen,
  steps_per_epoch = 50,
  epochs = 5,
  validation_steps = test_steps
)

plot(test)
```
From test result we see loss on test is .1669 which is less than baseline MAE of .45 which shows GRU model was able to reduce loss on partial data